\\pkg{casebase}: An Alternative Framework For Survival Analysis and Comparison of Event Rates

### 2021-10-25 analysis/paper/bhatnagar-turgeon-islam-hanley-saarela.Rmd

Rather than discretizing time, Hanley & Miettinen [-@hanley2009fitting] 
selected a discrete set of person-time coordinates (`person-moments') in 
continuous time from all observed follow-up experience constituting the 
study base. By doing so, they obtained a likelihood expression for the hazard 
function that is equivalent to that of logistic regression. More specifically, all person-moments when the event of interest 
occurred are selected as the case series, complemented by a randomly sampled 
base series of person-moments serving as controls. This approach allows 
flexible modeling of the hazard function by including time as a covariate 
(e.g. using splines or general additive models). Furthermore, time-dependent 
covariates can be modeled through interactions with time. In short, Hanley & 
Miettinen [-@hanley2009fitting] use the well-understood logistic regression 
for directly modeling the hazard function, without requiring a discrete-time model.

@@

```{r erspc-data-mutate, eval = eval_cs1, echo = FALSE}
data("ERSPC")
ERSPC$ScrArm <- factor(ERSPC$ScrArm, 
                       levels = c(0,1), 
                       labels = c("Control group", "Screening group"))
```

>>>>

Rather than discretizing time, Hanley & Miettinen [-@hanley2009fitting] 
selected a discrete set of person-time coordinates ("person-moments") in 
continuous time from all observed follow-up experience constituting the 
study base. By doing so, they obtained a likelihood expression for the hazard 
function that is equivalent to that of logistic regression. More specifically, all person-moments when the event of interest 
occurred are selected as the case series, complemented by a randomly sampled 
base series of person-moments serving as controls. This approach allows 
flexible modeling of the hazard function by including time as a covariate 
(e.g. using splines or general additive models). Furthermore, time-dependent 
covariates can be modeled through interactions with time. In short, Hanley & 
Miettinen [-@hanley2009fitting] use the well-understood logistic regression 
for directly modeling the hazard function, without requiring a discrete-time model.

@@

```{r erspc-data-mutate, eval = eval_cs1, echo = FALSE}
data("ERSPC")
# ERSPC$ScrArm was turned to factor in 0.10.1
if (utils::packageVersion("casebase") < package_version("0.10.1")) {
  ERSPC$ScrArm <- factor(ERSPC$ScrArm, 
                       levels = c(0,1), 
                       labels = c("Control group", "Screening group"))
}
```
